---
title: "Titanic Challange"
subtitle: "The famous data set from Kaggle"
author: "Alexander Nyberg"
date: '2021-04-01'
output: pdf_document
fontsize: 18pt
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(corrplot)
```

Load the data sets.

```{r include=FALSE}
titanic_train <- read_csv("../Data/train.csv")
titanic_test <- read_csv("../Data/test.csv")
gender_sub <- read_csv("../Data/gender_submission.csv")

# what sorts of people were more likely to survive?
```

## EDA

### Training Data Analysis

```{r}
glimpse(titanic_train)
summary(titanic_train)
```

Mean age for training set is 29.70 for the passengers, youngest being 0.42 and oldest being 80. We also indicate 177 NA's in age. 
We can see that 38% of the people in the training set survived the disaster.

Plotting the survival rate.

```{r}
SurvDeath <- titanic_train %>%
  select(Survived)
  
  ggplot() + 
    geom_bar(data = SurvDeath, aes(x = Survived))
```

```{r}
# Identifying the missing values
df_missValue <- data.frame(Key = character(1), Value = integer(1), Perc = integer(1))
for (i in 1:ncol(titanic_train)) {
  df_missValue <- rbind(df_missValue, 
                        c(colnames(titanic_train[, i]), 
                        sum(is.na(titanic_train[, i])), 
                        sum(is.na(titanic_train[, i]))/nrow(titanic_train)))
}
df_missValue <- df_missValue %>%
  mutate(Value = as.integer(Value), 
         Perc = as.double(Perc)) %>%
  filter(Value != 0) %>%
  arrange(desc(Value))

knitr::kable(df_missValue)

df_missValue %>%
  ggplot() + 
    geom_bar(aes(x = Key, y = Value), stat = 'identity') + 
    labs(x = "Variable", y = "Number of missing values", title = "Missing Values")
```

```{r}
# Percentage of missing values
df_missValuePerc <- data.frame(Key = character(1), Value = integer(1), Tot = integer(1))
for (i in 1:ncol(titanic_train)) {
  df_missValuePerc <- rbind(df_missValuePerc, 
                        c(colnames(titanic_train[, i]), 
                        sum(is.na(titanic_train[, i])), 
                        nrow(titanic_train[, i])))
}

df_missValuePerc %>%
  mutate(Value = as.integer(Value),
         Tot = as.integer(Tot)) %>%
  filter(Value != 0) %>%
  mutate(PercentageMissing = Value/Tot*100) %>%
  knitr::kable()
```

```{r warning=FALSE}
titanic_train %>%
  select(Survived, Sex, Age) %>%
  group_by(Age) %>%
  ggplot() + 
    geom_col(aes(x = Age, y = Survived, fill = Sex), width = 1)
```

Majority of survivals is women. 

```{r}
# Total % of females that survived
femalesAlive <- titanic_train %>%
  filter(Sex == "female" & Survived == 1)

totFemales <- titanic_train %>%
  filter(Sex == "female")

knitr::kable(nrow(femalesAlive)/nrow(totFemales), 
             col.names = "SurvRateFemale")
``` 

```{r}
# Total % males that survived
malesAlive <- titanic_train %>%
  filter(Sex == "male" & Survived == 1)

totMales <- titanic_train %>%
  filter(Sex == "male")

knitr::kable(nrow(malesAlive)/nrow(totMales),
             col.names = "SurvRateMale")
```

### Test Data Analysis

In the test data, we can see that we are missing some *Age* values again. These will be fixed with the mean of the training data set. Additionally, we have a missing value in the *Fare* column as well. 

```{r}
glimpse(titanic_test)
summary(titanic_test)
```

### Full Data Set

Firstly, we might consider the specific variables that actually might effect death probability. For now, we'll drop PassengerId, Name and Ticket. We'll also drop the Cabin column since 77% of the data is missing.

```{r message=FALSE}
fullData <- titanic_train %>%
  full_join(titanic_test) %>%
  select(-PassengerId, -Name, -Ticket, -Cabin)

knitr::kable(fullData[1:15, ])
```

Since we have two different outcomes in the *Sex* column, we can easily change these to numeric.

```{r}
for (i in 1:nrow(fullData)) {
  if (fullData[i, "Sex"] == "male") {
    fullData[i, "Sex"] <- "0"
  }
  else{
    fullData[i, "Sex"] <- "1"
  }
}

fullData <- fullData %>%
  mutate(Sex = as.double(Sex))
```


### Handling Missing Values 

With the missing data in age, a good choice could be to find the mean or the median and fill the missing values since we dont have that many missing values. Same goes for embarked.

```{r}
# Finding age mean
meanAge <- fullData %>% 
  select(Age) %>%
  summarise(mean(Age, na.rm = TRUE)) %>%
  round(digits = 1)
```

```{r}
# Fill meanAge in missing Age values
for (i in 1:nrow(fullData)) {
  if (is.na(fullData[i, "Age"]) == TRUE) {
    fullData[i, "Age"] <- meanAge
  }
}
```

```{r}
# Fixing embarked data
for (i in 1:nrow(fullData)) {
  if (is.na(fullData[i, "Embarked"]))
    fullData[i, "Embarked"] <- "S"
}
```

```{r}
# Fixing the missing data in Fare
meanFare <- fullData %>% 
  select(Fare) %>% 
  summarise(median(Fare, na.rm = T)) %>% 
  round(digits = 1)

for (i in 1:nrow(fullData)) {
  if (is.na(fullData[i, "Fare"]) == TRUE) {
    fullData[i, "Fare"] <- meanFare
  }
}
```

```{r}
# Double Checking no data missing
df_missValue <- data.frame(Key = character(1), Value = integer(1), Perc = integer(1))
for (i in 1:ncol(fullData)) {
  df_missValue <- rbind(df_missValue, 
                        c(colnames(fullData[, i]), 
                        sum(is.na(fullData[, i])), 
                        sum(is.na(fullData[, i]))/nrow(fullData)))
}
df_missValue <- df_missValue %>%
  mutate(Value = as.integer(Value), 
         Perc = as.double(Perc)) %>%
  filter(Value != 0) %>%
  arrange(desc(Value))

knitr::kable(df_missValue)
```

The data set is now complete and we are to predict the 418 test individuals.

### Prediction

```{r}
xTrain <- fullData %>%
  filter(!is.na(Survived))
xTest <- fullData %>%
  filter(is.na(Survived))
```

```{r}
tCorr <- cor(xTrain[,-8])

corrplot(tCorr, method = "number", type = "upper")
```

#### Linear Regression

```{r}
# Linear Regression Prediction
LR <- lm(Survived ~ . - Embarked, data = xTrain)
summary(LR)
```

```{r}
# Presenting the results
testSurv <- data.frame("SurvivedAlle" = round(predict(LR, xTest)))
testResult <- testSurv %>%
  mutate(PassengerId = titanic_test$PassengerId) %>% 
  select(PassengerId, SurvivedAlle)
row.names(testResult) <- NULL

knitr::kable(head(testResult))
```

```{r message=FALSE}
# Comparison Results 99% Accuracy. Answers not owned or created by myself.
compareData <- read_csv("../Data/submit.csv")
```

```{r}
compareDF <- testResult %>%
  left_join(compareData, by = "PassengerId") %>%
  rename("Survived99PercentAcc" = Survived)
knitr::kable(head(compareDF))
```

```{r}
identVector <- c()
for (i in 1:nrow(compareDF)) {
  if (compareDF[i, "SurvivedAlle"] == compareDF[i, "Survived99PercentAcc"]) {
    identVector <- append(identVector, T)
  }
  else{
    identVector <- append(identVector, F)
  }
}
compareDF$Alike <- identVector
```

```{r}
compareDF %>%
  group_by(Alike) %>% 
  count() %>%
  knitr::kable(col.names = c("Key", "Counts"))
```


